Investigating Overfitting in Convolutional Neural Networks

### 2025-05-23 paper_formatted.tex
added figures/ graphs, and started calculations section.

% Packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{float}
\usepackage{tikz}
\usepackage{setspace}
\usepackage{times}
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}

@@

    We trained $100$ convolutional models each on the same randomly selected small subset of 100 images
    from the Canadian Institute For Advanced Research - 10 dataset (CIFAR-10), which consists of 32x32 color images labeled with one one of ten classes (e.g., plane, boat, etc).
    Each model had the same architecture, the only difference between them being the number of parameters (\textit{see figure 1 below}).
    The training dataset was constructed by randomly selecting 10 of each class of image to ensure that it is representative of the entire dataset.
    The number of filters and neurons in each model were varied so that the model sizes are roughly uniform, ranging from approximately 1m - 50m parameters.
    The architecture and the ratio of layer sizes were also kept the same for each model, making sure the only difference between them is the number of parameters.
    On the initialization of each model in each set, the values of the parameters are randomly set, so each model can be seen as randomly
    selected from all possible models of their respective size and architecture before training, ensuring the experiment is statistically valid.
    We controlled the parameter count by multiplying the amount of filters for the Conv2D layers and the amount of neurons
    for the hidden Dense layers by a scalar factor, $n$.

    \noindent\textbf{Model Architecture}
    \begin{center}
        \begin{tikzpicture}[node distance=4mm and 0mm]
            \node[layer=gray!10]                     (input)  {Input $(32\times32\times3)$};
            \node[layer=blue!10,  below=of input]    (conv1)  {Conv2D: $32\cdot n$ filters, $3\times3$, ReLU};
            \node[layer=cyan!10,  below=of conv1]    (pool1)  {MaxPool: $2\times2$};
            \node[layer=blue!10,  below=of pool1]    (conv2)  {Conv2D: $64\cdot n$ filters, $3\times3$, ReLU};
            \node[layer=cyan!10,  below=of conv2]    (pool2)  {MaxPool: $2\times2$};
            \node[layer=orange!10,below=of pool2]    (flat)   {Flatten};
            \node[layer=green!15, below=of flat]    (dense1) {Dense: $128\cdot n$ neurons, ReLU};
            \node[layer=green!15, below=of dense1]  (dense2) {Dense: $64\cdot n$ neurons, ReLU};
            \node[layer=red!15,   below=of dense2]  (output) {Dense (Output): 10, Softmax};

            \draw[arrow] (input)  -- (conv1);
            \draw[arrow] (conv1)  -- (pool1);
            \draw[arrow] (pool1)  -- (conv2);
            \draw[arrow] (conv2)  -- (pool2);
            \draw[arrow] (pool2)  -- (flat);
            \draw[arrow] (flat)   -- (dense1);
            \draw[arrow] (dense1) -- (dense2);
            \draw[arrow] (dense2) -- (output);
        \end{tikzpicture}
    \end{center}

    \centerline{\textit{Figure 1: General Architecture of Our Models}}

@@

@@

    {The scatterplot of the data shows a clear linear trend, and the residual plot shows no remaining curved pattern. Therefore, the linearity condition is satisfied.}

@@

    \textbf{5.1.3 Normal}
    The dotplot of the residuals shows no strong skew and no outliers, suggesting the residuals are approximately normally distributed.

    \textbf{5.1.4 Equal Variance}
    The residual plot shows shows no noticeable $<$ or $>$ shape, so the equal variance condition is satisfied.

    \textbf{5.1.5 Random}
    The parameters of the model were randomized on initialization, so each model can be seen as randomly selected from the population
    of models of the same size and structure.

@@

>>>>

% Packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[labelfont=it, textfont=it]{caption}
\usepackage{float}
\usepackage{tikz}
\usepackage{setspace}
\usepackage{times}
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}

@@

    We trained $100$ convolutional models each on the same randomly selected small subset of 100 images
    from the Canadian Institute For Advanced Research - 10 dataset (CIFAR-10), which consists of 32x32 color images labeled with one one of ten classes (e.g., plane, boat, etc).
    Each model had the same architecture, the only difference between them being the number of parameters (\textit{see Figure 1 below}).
    The training dataset was constructed by randomly selecting 10 of each class of image to ensure that it is representative of the entire dataset.
    The number of filters and neurons in each model were varied so that the model sizes are roughly uniform, ranging from approximately 1m - 50m parameters.
    The architecture and the ratio of layer sizes were also kept the same for each model, making sure the only difference between them is the number of parameters.
    On the initialization of each model in each set, the values of the parameters are randomly set, so each model can be seen as randomly
    selected from all possible models of their respective size and architecture before training, ensuring the experiment is statistically valid.
    We controlled the parameter count by multiplying the amount of filters for the Conv2D layers and the amount of neurons
    for the hidden Dense layers by a scalar factor, $n$.

    \noindent\textbf{Model Architecture}
    \begin{figure}[H]
        \begin{center}
            \begin{tikzpicture}[node distance=4mm and 0mm]
                \node[layer=gray!10]                     (input)  {Input $(32\times32\times3)$};
                \node[layer=blue!10,  below=of input]    (conv1)  {Conv2D: $32\cdot n$ filters, $3\times3$, ReLU};
                \node[layer=cyan!10,  below=of conv1]    (pool1)  {MaxPool: $2\times2$};
                \node[layer=blue!10,  below=of pool1]    (conv2)  {Conv2D: $64\cdot n$ filters, $3\times3$, ReLU};
                \node[layer=cyan!10,  below=of conv2]    (pool2)  {MaxPool: $2\times2$};
                \node[layer=orange!10,below=of pool2]    (flat)   {Flatten};
                \node[layer=green!15, below=of flat]    (dense1) {Dense: $128\cdot n$ neurons, ReLU};
                \node[layer=green!15, below=of dense1]  (dense2) {Dense: $64\cdot n$ neurons, ReLU};
                \node[layer=red!15,   below=of dense2]  (output) {Dense (Output): 10, Softmax};

                \draw[arrow] (input)  -- (conv1);
                \draw[arrow] (conv1)  -- (pool1);
                \draw[arrow] (pool1)  -- (conv2);
                \draw[arrow] (conv2)  -- (pool2);
                \draw[arrow] (pool2)  -- (flat);
                \draw[arrow] (flat)   -- (dense1);
                \draw[arrow] (dense1) -- (dense2);
                \draw[arrow] (dense2) -- (output);
            \end{tikzpicture}
            \caption{General Architecture of Our Models}
        \end{center}
        \label{fig:architecture}
    \end{figure}

@@

    \begin{figure}[H]
        \centering
        \includegraphics[width=0.7\textwidth]{Images/Scatter}
        \caption{Scatterplot of Parameter Count vs. Difference in Train and Test Accuracy (Train - Test)}
        \label{fig:scatterplot}
    \end{figure}

    \begin{figure}[H]
        \centering
        \includegraphics[width=0.7\textwidth]{Images/Resid}
        \caption{Residual Plot (Parameter Count vs. Residuals)}
        \label{fig:residuals}
    \end{figure}

    \begin{figure}[H]
        \centering
        \includegraphics[width=0.7\textwidth]{Images/Resid Hist}
        \caption{Histogram of Residuals}
        \label{fig:residual_hist}
    \end{figure}

@@

    The scatterplot of the data (\textit{\autoref{fig:scatterplot}}) shows a clear linear trend, and the residual plot (\textit{\autoref{fig:residuals}}) shows no remaining curved pattern.
    Therefore, the linearity condition is satisfied.

@@

    \textbf{5.1.3 Normal}

    The histogram of the residuals (\textit{\autoref{fig:residual_hist}}) shows no strong skew and no obvious outliers,
    so the normality condition is met.

    \textbf{5.1.4 Equal Variance}

    The residual plot (\textit{\autoref{fig:residuals}}) shows shows no noticeable $<$ or $>$ shape, so the equal variance condition is satisfied.

    \textbf{5.1.5 Random}

    The parameters of the model were randomized on initialization, so each model can be seen as randomly selected from the population
    of models of the same size and structure.

@@

    We used the following equation to calculate the Standard Error $SE_\beta$ of the slope \beta:

    \begin{equation*}
        \centering
        SE_\beta = \frac{s}{s_x\sqrt{n - 1}}
    \end{equation*}
